---
title: "Data quality metrics"
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
bibliography: references.bib
csl: american-medical-association.csl
link-citations: yes
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float:
      collapsed: FALSE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, child = "js/back-to-top.js"}
```

```{css, echo = FALSE}
.warning {
  margin-bottom: 20px;
}
```

<br>

::: warning
<i class="fa fa-exclamation-circle" style="color: white"></i>&nbsp; THIS PAGE IS UNDER DEVELOPMENT
:::

Listed below are the data quality metrics that Gannet computes during data processing, signal fitting, and tissue segmentation.

## Linewidth

[TEXT]

## Signal-to-noise ratio (SNR)

The SNR of fitted signals is calculated as the amplitude of the given modeled signal divided by twice the standard deviation of the noise signal. To estimate noise, Gannet takes two independent segments of the OFF or DIFF spectrum (as appropriate to the modeled signal of interest) between 8–9 ppm and 9–10 ppm, and detrends them using a second-order polynomial function. The standard deviation of each detrended noise segment is then calculated. Detrending is required to remove baseline artifacts (often related to the water signal). The smaller of the two standard deviations is then used as the estimate of noise, which is then multiplied by 2.

Formulaically, this is defined as:

$$
SNR_{signal} = \frac{A_{signal}}{2*\mathrm{std}(noise)}
$$

where:

| <u>Parameter</u> | <u>Definition</u> |
| :- | :--------- |
| $A_{signal}$ | Signal model amplitude |
| $noise$ | Detrended noise signal between either 8–9 or 9–10 ppm (whichever is lower) in the appropriate spectrum (i.e., either the OFF or DIFF spectrum) |

## Frequency offsets (frequency drift/motion)

[TEXT]

## Fit error

When fitting signal functions to metabolite peaks, Gannet will also estimate the error of the fit. This is defined as the standard deviation of the residuals of the signal model fit divided by the signal model amplitude. For metabolite peak fits, this is:

$$
\epsilon_{metab} = 100\frac{\mathrm{std}(resid_{metab})}{A_{metab}}
$$

where:

| <u>Parameter</u> | <u>Definition</u> |
| :- | :--------- |
| $resid_{metab}$ | Signal model fit residuals |
| $A_{metab}$ | Signal model amplitude |

Similarly, for reference signal fits, this is:

$$
\epsilon_{ref} = 100\frac{\mathrm{std}(resid_{ref})}{A_{ref}}
$$

Since all metabolites are normalized to a reference signal (Cr or unsuppressed water), the fit error that really should be considered (and reported) is the combined error of the metabolite and reference signal model fits, which add up in quadrature. Formulaically, this is defined as:

$$
\epsilon_{metab,ref} = \sqrt{\epsilon_{metab}^2 + \epsilon_{ref}^2}
$$

## Tissue segmentation

[TEXT]



