---
title: "Data quality metrics"
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
bibliography: references.bib
csl: american-medical-association.csl
link-citations: yes
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float:
      collapsed: FALSE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, child = "js/back-to-top.js"}
```

```{css, echo = FALSE}
.warning {
  margin-bottom: 20px;
}
```

<br>

::: warning
<i class="fa fa-exclamation-circle" style="color: white"></i>&nbsp; THIS PAGE IS UNDER DEVELOPMENT
:::

Listed below are the preprocessing routines that are run in `GannetLoad.m`. Whether a particular routine or subroutine is run will depend on the format of the inputted data and the options set in `GannetPreInitialise.m`.

## Linewidth

[Text]

## Signal-to-noise ratio (SNR)

[Text]

## Fit error

When fitting signal functions to metabolite peaks, Gannet will also estimate the error in the fit. This is defined as the signal model amplitude divided by the standard deviation of the model fit residuals:

$$
\epsilon_{metab} = A_{metab}/std(resid)
$$

where:

| <u>Parameter</u> | <u>Definition</u> |
| :- | :--------- |
| $A_{metab}$ | Signal model amplitude |
| $resid$ | Signal model fit residuals |

Since all metabolite are normalized to a reference signal (Cr or unsuppressed water), the fit error that really should be considered (and reported) is the combined error of the metabolite and reference signal model fits, which add up in quadrature. Formulaically, this is defined as:

$$
\epsilon_{metab,ref} = \sqrt{\epsilon_{metab}^2 + \epsilon_{ref}^2}
$$

## Tissue segmentation

[Text]





