---
title: "FAQ"
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: TRUE
    toc_depth: 3
    toc_float:
      collapsed: FALSE
---

<style>
body .main-container {
  max-width: 1200px;
}
</style>

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, child = "js/back-to-top.js"}
```

<br>

### General

#### **Can Gannet process GSH-edited data?**

Yes, Gannet can process GABA-, GSH-, Lac-, and EtOH-edited MEGA-PRESS data, as well as GABA-/GSH-, Lac-/GSH-, and EtOH-/GABA-/GSH-edited HERMES and HERCULES data. (Gannet does not support the fitting of HERCULES data.) Simply set the appropriate `target` option(s) in `GannetPreInitialise.m`.

#### **Can Gannet process phantom data?**

Yes, of the edited data described above. Make sure to set `phantom` to `1` under 'Flags' in `GannetPreInitialise.m`. If you want to fit your data, run `GannetFitPhantom(MRS_struct);` instead of `GannetFit(MRS_struct);`. (Fitting of HERCULES phantom data is not supported.)


### Output

#### **[BLANK]**


### Errors

#### **[BLANK]**






