---
title: "Output structure attributes"
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float:
      collapsed: FALSE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, child = "js/back-to-top.js"}
```

```{css, echo = FALSE}
.info {
  margin-bottom: 20px;
}
```

<br>

::: info
<i class="fa fa-info-circle" style="color: white"></i>&nbsp; The fields listed on this page are applicable to release 3.3.0 and are not necessarily correct for other versions of Gannet.
:::

All output from Gannet analyses are saved in a structure (e.g., `MRS_struct`, or whatever name the user has chosen) that contains many fields and subfields. Listed below are lists of these fields and their descriptions. Note that not possible field that Gannet can generate during an analysis pipeline is shown for the sake of brevity.

## MRS_struct

| <u>Field</u> | <u>Description</u> |
|:------------------|:----------------------------------------------------|
| `version` | Structure containing release numbers of Gannet and the Gannet modules |
| `ii` | During analysis, the current file being loaded; after analysis, the number of files that were loaded |
| `metabfile` | Cell array containing the inputted metabolite data filenames |
| `waterfile` | Cell array containing the inputted water reference data filenames |

## MRS_struct.p

| <u>Field</u> | <u>Description</u> |
|:------------------|:----------------------------------------------------|
`target` | The metabolite(s) targeted by the editing pulses |
`seqorig` | Origin of the Philips MEGA-PRESS or GE HERMES sequences |
`LB` | Exponential line-broadening applied (Hz) |
`water_ECC` | Whether eddy-current correction was applied to the water data |
`metab_ECC` | Whether eddy-current correction was applied to the metabolite data |
`water_removal` | Whether the residual water signal in DIFF spectrum was removed using HSVD |
`alignment` | Choice of shot-to-shot frequency-and-phase correction |
`use_prealign_ref` | Whether `RobustSpecReg` used the averaged pre-aligned subspectra as references to align the averaged post-aligned subspectra |
`fit_resid_water` | Whether the residual water signal in the OFF spectrum was fitted to calculate a water suppression factor |
`weighted_averaging` | Whether weighted signal averaging was implemented |
`HERMES` | Whether the data are HERMES data |
`HERCULES` | Whether the data are HERCULES data | 
`PRIAM` | Whether the data are PRIAM data |
`phantom` | Whether the data are phantom data |
`join` | Whether the input files were joined |
`mat` | Whether the output structure was saved as MAT-file |
`csv` | Whether a CSV file of useful data was created |
`append` | Whether the PDF outputs were appended into single files |
`hide` | Whether the output figures were hidden |
`vendor` | Scanner vendor |
`reference` | Concentrations will be calculated relative to this; if a water reference is provided, concentrations will be calculated relative to both water and Cr, otherwise they will be calculated relative to Cr |
`numScans` | Total number of scans in the batch |
`numFilesPerScan` | When input files are joined, the number of files joined per scan |
`npoints` | Number of data points of each input file |
`nrows` | Number of data frames stored in each input water reference file |
`nrows_water` | Number of data frames stored in each input file |
`Navg` | Total number of averages collected at acquisition for each input file |
`Nwateravg` | Total number of averages collected at acquisition for each input water reference file |
`TR` | Repetition time (ms) |
`TE` | Echo time (ms) |
`LarmorFreq` | Larmor frequency (MHz) |
`sw` | Spectral width (Hz) |
`voxdim` | Array of voxel dimensions (mm) |
`voxoff` | Array of voxel position offset (mm) |
`voxang` | Array of voxel angulation (deg) |
`ZeroFillTo` | Processed spectra are zero-filled to this many data points |
`zf` | Zero-fill factor |
`dt` | Dwell time (s) |
`SpecRes` | Spectral resolution of the raw data (Hz/point) |
`SpecResNominal` | Nominal spectral resolution of the processed data (Hz/point) |
`Tacq` | Acquisition time (s) |

## MRS_struct.fids

| <u>Field</u> | <u>Description</u> |
|:------------------|:----------------------------------------------------|
| `data` | Raw time-domain metabolite data (dimensions: `npoints` $\times$ `nrows`) of the last loaded dataset |
| `data_water` | Raw time-domain water reference data (dimensions: `npoints` $\times$ `nrows_water`) of the last loaded dataset |
| `ON_OFF` | Array of the editing order of the last loaded dataset |
| `data_align` | Raw time-domain metabolite data of the last loaded dataset with frequency-and-phase correction applied |

## MRS_struct.spec

| <u>Field</u> | <u>Description</u> |
|:------------------|:----------------------------------------------------|
| `freq` | Frequency axis (ppm), calculated from the last loaded dataset |
| `F0freq` | Cell array of the observed frequency of either residual water signal or the 3 ppm Cr signal for each average |
| `F0freq2` | Cell array of the observed frequency of the 3 ppm Cr signal for each average; this is used in `RobustSpectralRegistration.m` as the starting values for frequency during optimization |
| `AllFramesFT` | Spectra of the last loaded dataset without frequency-and-phase correction applied |
| `AllFramesFTrealign` | Spectra of the last loaded dataset with frequency-and-phase correction applied |

## MRS_struct.out

| <u>Field</u> | <u>Description</u> |
|:------------------|:----------------------------------------------------|
| `AvgDeltaF0` | A metric of the amount of frequency offset observed during the acquisition, calculated as the mean difference between the observed frequency of the residual water signal (or the 3 ppm Cr signal) in the pre-frequency-corrected subspectra and the nominal water frequency at 4.68 ppm (or the nominal 3 ppm Cr signal if HERMES or GSH editing) |
| `SpecReg` | Structure containing output from spectral registration |
| `reject` | Cell array of binary integers indicating whether an individual transient should be rejected during signal averaging |
| `QA` | Metrics calculated to quantify the performance of tissue segmentation |

## MRS_struct.mask

| <u>Field</u> | <u>Description</u> |
|:------------------|:----------------------------------------------------|
| `outfile` | Cell array of the output filenames of the voxel masks |
| `img` | Cell array containing three-dimensional arrays representing the three orthogonal slices of the T1 images |
| `T1image` | Cell array of the input T1 image filenames |
| `img_montage` | Cell array containing three-dimensional arrays representing a montage of axial slices of the T1 images overlaid by the tissue-segmented voxel masks |



