---
title: "Output structure attributes"
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float:
      collapsed: FALSE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, child = "js/back-to-top.js"}
```

<br>

::: info
<i class="fa fa-info-circle" style="color: white"></i>&nbsp; The fields listed on this page are applicable to release 3.3.0 and are not necessarily correct for other versions of Gannet.
:::

## MRS_struct

| <u>Field</u> | <u>Description</u> |
| :- | :--------- |
| `version` | Release numbers of Gannet and the Gannet modules |
| `ii` | During analysis, the current file being loaded; after analysis, the number of files that were loaded |
| `metabfile` | String array containing the inputted metabolite data filenames |
| `waterfile` | String array containing the inputted water reference data filenames |

## MRS_struct.p

| <u>Field</u> | <u>Description</u> |
| :- | :--------- |
`Reference_compound` | Concentrations will be calculated relative to this; if a water reference is provided, concentrations will be calculated relative to both water and Cr, otherwise they will be calculated relative to Cr |
`target` | 
`ONOFForder` | 
`seqorig` | 
`LB` | 
`water_phase_correction` |
`data_phase_correction` |
`water_removal` |
`AlignTo` |
`Vox` | 
`FitResidWater` |
`HERMES` |
`HERCULES` | 
`PRIAM` |
`phantom` |
`mat` |
`sdat` |
`csv` |
`vendor` | Scanner vendor |
`spar_string` |
`numscans` | Total number of scans in the batch |
`npoints` | Number of data points |
`nrows` |
`Navg` | Total number of averages |
`TR` | Repetition time (ms) |
`TE` | Echo time (m) |
`LarmorFreq` | Larmor frequency (kHz) |
`sw` | Spectral width (Hz) |
`voxdim` | Voxel dimensions (mm) |
`voxoff` | Voxel position offset (mm) |
`voxang` | Voxel angulation (deg) |
`Nwateravg` | Total number of water averages |
`ZeroFillTo` | Processed spectra are zero-filled to this many data points |
`zf` | Zero-filling factor |
`df` | 
`SpecRes` | Spectral resolution of the raw data |
`SpecResNominal` | Nominal spectral resolution of the processed data |
`Tacq` | Acquisition time |

## MRS_struct.fids

| <u>Field</u> | <u>Description</u> |
| :- | :--------- |
| `data` | Raw time-domain metabolite data (dimensions: `npoints` $\times$ `nrows`) of the last loaded dataset |
| `data_water` |
| `ON_OFF` |
| `data_align` |
| `vox1` |



