---
title: "Output structure attributes"
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float:
      collapsed: FALSE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, child = "js/back-to-top.js"}
```

```{css, echo = FALSE}
.info {
  margin-bottom: 20px;
}
```

<br>

::: info
<i class="fa fa-info-circle" style="color: white"></i>&nbsp; The fields listed on this page are applicable to release 3.3.0 and are not necessarily correct for other versions of Gannet.
:::

All the output from Gannet anaylses are saved in a structure (e.g., `MRS_struct` or whatever name the user has chosen) that contains many fields and subfields. Listed below are lists of these fields and their descriptions. Note that not every possible field that Gannet can generate is shown for the sake of brevity.

## MRS_struct

| <u>Field</u> | <u>Description</u> |
| :- | :--------- |
| `version` | Structure containing release numbers of Gannet and the Gannet modules |
| `ii` | During analysis, the current file being loaded; after analysis, the number of files that were loaded |
| `metabfile` | Cell array containing the inputted metabolite data filenames |
| `waterfile` | Cell array containing the inputted water reference data filenames |

## MRS_struct.p

| <u>Field</u> | <u>Description</u> |
| :- | :--------- |
`target` | The metabolite(s) targeted by the editing pulses |
`seqorig` | 
`LB` | 
`water_ECC` |
`metab_ECC` |
`water_removal` |
`alignment` |
`vox` | 
`fit_resid_water` |
`weighted_averaging` |
`HERMES` |
`HERCULES` | 
`PRIAM` |
`phantom` |
`mat` |
`sdat` |
`csv` |
`vendor` | Scanner vendor |
`reference` | Concentrations will be calculated relative to this; if a water reference is provided, concentrations will be calculated relative to both water and Cr, otherwise they will be calculated relative to Cr |
`numScans` | Total number of scans in the batch |
`npoints` | Number of data points |
`nrows` |
`Navg` | Total number of averages |
`TR` | Repetition time (ms) |
`TE` | Echo time (m) |
`LarmorFreq` | Larmor frequency (kHz) |
`sw` | Spectral width (Hz) |
`voxdim` | Voxel dimensions (mm) |
`voxoff` | Voxel position offset (mm) |
`voxang` | Voxel angulation (deg) |
`Nwateravg` | Total number of water averages |
`ZeroFillTo` | Processed spectra are zero-filled to this many data points |
`zf` | Zero-filling factor |
`df` | 
`SpecRes` | Spectral resolution of the raw data |
`SpecResNominal` | Nominal spectral resolution of the processed data |
`Tacq` | Acquisition time |

## MRS_struct.fids

| <u>Field</u> | <u>Description</u> |
| :- | :--------- |
| `data` | Raw time-domain metabolite data (dimensions: `npoints` $\times$ `nrows`) of the last loaded dataset |
| `data_water` |
| `ON_OFF` |
| `data_align` |
| `vox1` |

## MRS_struct.spec

| <u>Field</u> | <u>Description</u> |
| :- | :--------- |
| `data` | Raw time-domain metabolite data (dimensions: `npoints` $\times$ `nrows`) of the last loaded dataset |
| `data_water` |
| `ON_OFF` |
| `data_align` |
| `vox1` |

## MRS_struct.out

| <u>Field</u> | <u>Description</u> |
| :- | :--------- |
| `AvgDeltaF0` | A metric of the degree of frequency offsets during the acquisition relative to the nominal water frequency (4.68 ppm) or the 3 ppm Cr signal (if HERMES or GSH editing) |
| `SpecReg` | Structure containing output from spectral registration |
| `reject` | Cell array of binary integers indicating whether an individual transient should be rejected during signal averaging |
| `QA` | Metrics calculated to quantify the performance of tissue segmentation |

## MRS_struct.mask

| <u>Field</u> | <u>Description</u> |
| :- | :--------- |
| `outfile` | Cell array of the output filenames of the voxel masks |
| `img` | Cell array containing three-dimensional arrays representing the three orthogonal slices of the T1 images |
| `T1image` | Cell array of the input T1 image filenames |
| `img_montage` | Cell array containing three-dimensional arrays representing a montage of axial slices of the T1 images overlaid by the tissue-segmented voxel masks |



