---
title: "Preprocessing"
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
bibliography: references.bib
csl: american-medical-association.csl
link-citations: yes
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float:
      collapsed: FALSE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, child = "js/back-to-top.js"}
```

```{css, echo = FALSE}
table {
  margin: auto;
}

table thead th {
  border-bottom: 1px solid #ddd;
}

th, td {
  padding: 5px;
}
```

<br>

::: warning
<i class="fa fa-exclamation-circle" style="color: white"></i>&nbsp; THIS PAGE IS UNDER DEVELOPMENT
:::

<br>

Listed below are the preprocessing routines that are run in `GannetLoad.m`. Whether a particular routine or subroutine is run will depend on the format of the inputted data and the options set in `GannetPreInitialise.m`.

## RF coil combination

Certain raw MRS data formats store data without coil combination; specifically, GE P-file (.7), NIfTI-MRS (if the source data were raw), Philips .raw, and Siemens TWIX (.dat) data. Gannet uses generalized least squares [@An2013] to optimally combine the signal from the multiple RF channels. If water files are provided, these data will be used as references for signal weighting and phasing of the coil data.

## Eddy-current correction

::: info
<i class="fa fa-info-circle" style="color: white"></i>&nbsp; Eddy-current correction can only be applied if water data are provided.
:::

In `GannetPreInitialise.m`, users have the option to apply eddy-current correction (ECC) to metabolite and water data. If applied, Gannet uses the method described by Klose (1990) [@Klose1990]. The code for the ECC routine can be found in `EddyCurrentCorrection.m`.

## Line-broadending (apodization)

FID data are multiplied by a time-varying exponential weighting function where the weighting constant is set in `GannetPreInitialise.m` (3 Hz is the default).

## Frequency and phase alignment

During acquisition, spectral data are affected by frequency and phase offsets as a results of biophysical, electrical, mechanical, and participant factors. Gannet has several algorithms to correct for these errors during preprocessing. Users can choose which method to use in `GannetPreInitialise.m`. The methods are:

| <u>Method</u> | <u>Option</u> | <u>Description</u> |
| :-------- | :---- | :--------- |
| Robust spectral registration [@Mikkelsen2020] (default) | `RobustSpecReg` | A method based on spectral registration [@Near2015] that is robust against spectral distortions such as unstable residual water and lipid contamination.  |
| Multi-step frequency and phase correction [@Mikkelsen2018] | `SpecRegHERMES` | A method originally developed to align multiplexed edited HERMES data. This approach is also based on spectral registration. |

## Phase correction

It is common for unprocessed spectra to be out of phase. Gannet applies a global zero-order phase correction to all transients by fitting the real-valued 3 ppm Cr and 3.2 ppm Cho signals in the frequency domain.

## Signal averaging

Gannet provides two methods for averaging individual transients (selected in `GannetPreInitialise.m`): arithmetic averaging (with outlier rejection) and weighted averaging (the default). The code for the signal averaging routines can be found in `SignalAveraging.m`.

### Conventional averaging

Conventional (or arithmetic) averaging is straightforward. All sequentially acquired pairs of subspectra (e.g., all edit-ON and edit-OFF subspectra) are averaged using the arithmetic mean: $\bar{x} = \frac{1}{n}\sum_{i=1}^nx_i$.

Note that before the arithmetic averaging of subspectra, individual transients are excluded based on the outlier rejection algorithm used during frequency and phase alignment.

### Weighted averaging

Weighted averaging down-weights individual difference subspectra that are corrupted by signal artifacts — this is an important distinction from traditional signal averaging. First, the difference between sequentially acquired pairs (e.g., all edit-ON and edit-OFF subspectra) is calculated. A similarity matrix $\mathbf{D}\in\mathbb{R}^{P{\times}P}$ is obtained by calculating the mean squared error between each real-valued difference subspectrum $p$ and every other real-valued difference subspectrum (in the range 1.8 to 3.4 ppm). A similarity metric $d_{p}$ is calculated as the column-wise median of $\mathbf{D}$. Normalized weights $w_{p}$ are then derived, $w_{p} = d^{-2}_p/\sum{d^{-2}_p}$, and applied to the difference pairs before summation.

<br>

### References



