---
title: "Preprocessing"
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
bibliography: references.bib
csl: american-medical-association.csl
link-citations: yes
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float:
      collapsed: FALSE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, child = "js/back-to-top.js"}
```

```{css, echo = FALSE}
table {
  margin: auto;
}

table thead th {
  border-bottom: 1px solid #ddd;
}

th, td {
  padding: 5px;
}
```

<br>

Listed below are the preprocessing routines that are run in `GannetLoad.m`. Whether a particular routine or subroutine is run will depend on the format of the inputted data and the options set in `GannetPreInitialise.m`.

## RF coil combination

Certain raw MRS data formats store data without coil combination; specifically, GE P-file (.7), NIfTI-MRS (if the source data were raw), Philips .raw, and Siemens TWIX (.dat) data. Gannet uses generalized least squares [@An2013] to optimally combine the signal from the multiple RF channels. If water files are provided, these data will be used as references for signal weighting and phasing of the coil data.

## Eddy-current correction

In `GannetPreInitialise.m`, users have the option to apply eddy-current correction (ECC) to metabolite and water data. If applied, Gannet uses the method described by Klose (1990) [@Klose1990]. The ECC routine can be round in `EddyCurrentCorrection.m`.

::: info
<i class="fa fa-info-circle" style="color: white"></i>&nbsp; ECC can only be applied if water data are provided.
:::

## Line-broadending (apodization)

FID data are line-broadened using a time-varying exponential weighting function where the weighting constant set in `GannetPreInitialise.m` (3 Hz is the default).

## Frequency and phase alignment

During acquisition, spectral data are affected by frequency and phase offsets as a results of biophysical, electrical, mechanical, and participant factors. Gannet has several algorithms to correct for these errors during preprocessing. Users can choose which method to use in `GannetPreInitialise.m`. The methods are:

| <u>Method</u> | <u>Option</u> | <u>Description</u> |
| :-------- | :---- | :--------- |
| Robust spectral registration [@Mikkelsen2020] (default) | `RobustSpecReg` | A method based on spectral registration [@Near2015] that is robust against spectral distortions such as unstable residual water and lipid contamination.  |
| Multi-step frequency and phase correction [@Mikkelsen2018] | `SpecRegHERMES` | A method originally developed to align multiplexed edited HERMES data. This approach is also based on spectral registration. |

## Phase correction

It is common for unprocessed spectra to be out of phase. Gannet applies a global zero-order phase correction to all transients by fitting the 3 ppm Cr and 3.2 ppm Cho peaks in the.

## Signal averaging

Gannet provides two methods for averaging individual transients: arithmetic averaging (with outlier rejection) and weighted averaging (the default).

### Arithmetic averaging

Text.

### Weighted averaging

Text.

<br>

### References








