---
title: "Release notes"
date: "Last updated: `r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: TRUE
    toc_depth: 2
    toc_float:
      collapsed: FALSE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, child = "js/back-to-top.js"}
```

<br>

This page lists the stable releases of Gannet since v3.1.3. All releases can be found on GitHub <a href="https://github.com/markmikkelsen/Gannet/releases" target="_blank">here</a> and <a href="https://github.com/richardedden/Gannet3.1/releases" target="_blank">here</a>. The development version of Gannet can be found <a href="https://github.com/markmikkelsen/Gannet/tree/dev" target="_blank">here</a>. For a more comprehensive list of changes to the current main version, see the commits logged on GitHub <a href="https://github.com/markmikkelsen/Gannet/commits/main" target="_blank">here</a>.

## 3.5.2
<span style="color: grey;">2026-01-30</span>

### What's changed

* Added dynamic font size adjustment for R2025a+
* Refactored segmentation and QA metrics in `Seg.m`
* Consolidated dependency update workflows
* Fixed spacing in MRIQC metrics comment
* Added linewidth parameter to `PaperPlot.m`
* Removed blank issue template
* Added more labels to the image montage in GannetQuantify output figure
* Updated `README.md`: logo path, compatibility, and contributors
* Updated `CoRegStandAlone.m` settings and added normalization to MNI space
* Added workflow to automerge Dependabot PRs
* Updated dependencies - bids-matlab, dic2nii, etc.
* Bumped actions/checkout from 5 to 6
* Bumped peter-evans/create-pull-request from 7 to 8

**Full changelog**: <a href="**Full Changelog**: https://github.com/markmikkelsen/Gannet/compare/v3.5.1...v3.5.2" target="_blank">`v3.5.1...v3.5.2`</a>

## 3.5.1
<span style="color: grey;">2025-09-13</span>

### What's changed

* Set figure theme to light for MATLAB R2025a and newer
* Minor bug fixes for MATLAB R2025a

**Full changelog**: <a href="https://github.com/markmikkelsen/Gannet/compare/v3.5.0...v3.5.1" target="_blank">`v3.5.0...v3.5.1`</a>

## 3.5.0
<span style="color: grey;">2025-09-11</span>

### New features

- Gannet is now compatible with <a href="https://bids-specification.readthedocs.io/en/stable/" target="_blank">BIDS</a>
  - [`bids-matlab`](https://bids-matlab.readthedocs.io/en/main/), a MATLAB toolkit for interfacing with BIDS datasets, is embedded in Gannet, and users can now input BIDS datasets directly into `GannetLoad.m` (see [documentation](https://markmikkelsen.github.io/Gannet-docs/getting-started.html#compatibility-with-bids) for usage instructions)
- Added image QA metrics based on <a href="https://doi.org/10.1371/journal.pone.0184661" target="_blank">MRIQC</a>
  - `GannetSegment.m` now calculates and saves the following QA metrics for each anatomical image:
    - Signal-to-noise ratio (SNR) of GM, WM, CSF, and the whole brain (two variants: one using the median absolute deviation of the air background based on <a href="https://doi.org/10.1002/jmri.20969" target="_blank">Dietrich et al. (2007)</a>, and one using the standard deviation of the noise within the tissue class)
    - Contrast-to-noise ratio (CNR) between GM and WM
    - Coefficient of variation (CV) of GM and WM
    - Coefficient of joint variation (CJV) between GM and WM
    - Entropy-focus criterion (EFC)
    - Foreground-background energy ratio (FBER)
    - WM to maximum intensity ratio (WM2MAX): the median intensity within WM over the 95% percentile of the full intensity distribution, that captures the existence of long tails due to hyper-intensity of the carotid vessels and fat
  - These metrics are saved in `MRS_struct.out.QA`
- Added GitHub Actions for continuous integration with `bids-matlab`, `dicm2nii`, and `export_fig` to ensure compatibility with the latest versions of these dependencies
- Added `dependabot.yml` to automatically check for updates to GitHub Actions
- Water-unfiltered (but aligned) spectra are now saved in `MRS_struct`
- All version strings (e.g., `MRS_struct.version.*`) are now stored under `MRS_struct.info.version.*`, and datetime stamps for processing steps are added as `MRS_struct.info.datetime.*`

### Major changes

- When calculating GM, WM, and CSF fractions in `GannetSegment.m`, the probabilistic tissue maps are now thresholded at 0.9 before calculating the fractions. This should improve the accuracy of tissue fraction estimates

### Minor changes

- Modified `SiemensTWIXRead.m` to use iceParam for `pointsBeforeEcho` in MEGA-sLASER, aligning with MEGA-PRESS handling for consistency
- Changed the naming convention for water residual fit results from `ResidWater` to `resid_water` in `GannetFit.m` for consistency
- Improved input argument validation and error messages, including checks for cell array inputs and file existence, with clearer error output for missing files
- Updated `export_fig` to v3.52
- Updated `dicm2nii` to v2025.08.21
- MANY COSMETIC CHANGES to improve code readability and maintainability

**Full changelog**: <a href="https://github.com/markmikkelsen/Gannet/compare/v3.4.0...v3.5.0" target="_blank">`v3.4.0...v3.5.0`</a>

## 3.4.0
<span style="color: grey;">2025-03-30</span>

### New features

- Added support for fitting the 2.34 ppm Glu peak in the SUM spectrum
- Added `DeIdentifyDisclaimer.m`, which prints a disclaimer that the developers provide no guarantee that the de-identification functions will remove all human participants' identifying information in the data header of files
- The RF coil combination used is now stored in `MRS_struct.p.RFCoilCombination` (note: this only applies to raw data files that are not coil-combined during export from the scanner)
- Data load time (`MRS_struct.loadtime`) is now stored when `GannetLoad.m` is run
- Re-added the option to choose the editing pulse order in `GannetPreInitialise.m`: `MRS_struct.p.ON_OFF_order`
- MRS voxel masks can now be normalized to MNI152 space; a flag in `GannetPreInitiliase.m` has been added to do so (note that this is only performed if GannetSegment has been run as the forward deformation map needed for MNI normalization is created during SPM12's segmentation routine)
- Added `VoxelMaskOverlap.m`, a function that will create a voxel mask that indicates the degree of overlap of all datasets run in a batch
- Added `ThreeLorentzModel.m`
- Added `VoigtModel.m`

### Minor changes

- An "intro" message is printed whenever `GannetLoad.m` is run: the version number and a link to the software documentation website is displayed
- Co-registration figure outputs now show the anatomical directions of the displayed MR image (i.e., anterior/posterior [A/P], left/right [L/R], and superior/inferior [S/I])
- Replaced file separators with `filesep` so that they are specific to users' operating system
- `voxoff` is now parsed in `GERead.m` instead of in `GannetMask_GE.m` and `GannetMask_GE_nii.m`
- Minor changes to rounding in `ExportToCSV.m`
- Renamed `FitPeaksByFrames2.m` to `FitPeaksByFrames.m`
- Updated `export_fig` to v3.48
- Updated `dicm2nii` to v2024.08.21
- MANY COSMETIC CHANGES AND CLEANUP OF OLD UNUSED CODE

### Bug fixes

- Updated `AlignSubSpectra.m` to include Lac and GSH editing
- Updated `AlignUsingPeak.m` to fit the indexing approach of Gannet 3
- Fixed issue with Siemens .RDA files in `CoRegStandAlone.m`
- Added `MRS_struct.p.seqorig` to `CoRegStandAlone.m`
- Minor fix in `GannetMask_NIfTI.m` for how voxel offset is parsed from the NIfTI file
- Fixed an issue with `GannetFit.m` when users choose to fit a specific metabolite
- Corrected a bug in `PaperPlot.m` for plotting model fits for GABA+Glx
- Fixed an issue in `AlignUsingPeak.m` when running more than one file in a batch

## 3.3.2
<span style="color: grey;">2023-08-02</span>

### Major changes

- Rewrote `SiemensRead.m` to reduce code length

### Minor changes

- Updated `DICOMRead.m`, including to support Siemens XA30 data
- Updated `SpectralRegistrationHERMES.m`
  - SNR limit is now estimated as in `RobustSpectralRegistration.m`
  - Bug fix for indexing for output saved from maximum likelihood estimation
- Updated `dicm2nii` to version 2023.03.16
- Updated `export_fig` to v3.40
- Added Gannet documentation website URL to error dialog box
- Changed the way path names are printed in error dialog box
- Turned off iteration limit warning in `FitChoCr.m`
- Updated `UpdateGannet.m` to ensure all sub-directories are added to the search path after updating
- Added issue templates (in `.github/ISSUE_TEMPLATE`) for blank issues, bug reports, and feature requests
- Refined alignment of text/logos in output figures
  - The units used for displaying textual results are now normalized to fit to any subplot dimension
  - The textual results in the `GannetCoRegister`, `GannetSegment`, `CoReg`, and `Seg` output figures are now correctly centered
- The weighted signal averaging method used is now saved in the structure output
- The weighting factors calculated when applying weighted signal averaging are also now saved in the output
- Updated `SavePDF.m` such that when appending PDFs using `export_fig`, the figure size will now be equivalent in pixels on Windows and Mac systems
- Updated `NIfTIMRSRead.m`
- Updated `README.md`
- Updated `GannetFitPhantom.m`
- Removed redundant global frequency shift in `RobustSpectralRegistration.m`
- The z-standardized MSEs calculated in the spectral registration subroutines are now saved in the structure output
- Updated `GannetVersion.m` to also output version numbers of Gannet modules
- A warning is now printed when users process Siemens .rda data encouraging them to instead use TWIX (.dat) data
- Added support for GE P-file header revision 30 (<a href="https://github.com/markmikkelsen/Gannet/issues/28" target="_blank">#28</a>)
- MANY COSMETIC CHANGES

### Bug fixes

- Bug fix for special character `\` in Windows file path names printed in the error dialog box
- Fixed bug in `GannetMask_NIfTI.m` where the co-registered MRS voxel was incorrectly flipped in the `GannetCoRegister` output figure
- Fixed bug in `GannetQuantify.m` where the alpha value was incorrectly being printed multiple times when multiple were run in batch
- Added fix in `VersionCheck.m` so that when users have limited or restricted internet access, an error is avoided (<a href="https://github.com/markmikkelsen/Gannet/pull/27" target="_blank">#27</a>) - thanks <a href="https://github.com/alexcraven" target="_blank">@alexcraven</a>!

## 3.3.1
<span style="color: grey;">2023-03-14</span>

### New features

- Compressed NIfTI files (\*.nii.gz) are now fully supported
- Gannet now fully quantifies metabolite levels for Cr, Cho, and NAA (estimated from the OFF
  spectrum)

### Major changes

- Moved local signal model functions out of `GannetFit.m` and into their own files:
  - `BaselineModel.m`
  - `CalcIU.m`
  - `DoubleGaussModel.m`
  - `EtOHModel.m`
  - `GABAGlxModel.m`
  - `GaussModel.m`
  - `LorentzGaussModel.m`
  - `LorentzGaussModelP.m`
- Added `GetFullPath.m`; Gannet now will find the full path of input files by default
- Renamed some of the data-loading functions

### Minor changes

- Added a check in `GannetQuantify.m` for datasets without water references; `GannetQuantify.m` will
  not run if no water references are present
- Added a check in `GannetCoRegister.m` to ensure all structural image files can be found
- Updated `export_fig` to v3.33
- Updated `dicm2nii` to version 2023.03.08
- Updated `LICENSE` and `README.md`
- Updated `ExportToCSV.m`
- Updated `CoRegStandAlone` functions
- Replaced use of `1i` with `complex()`
- Removed unused variable `T1max`
- Removed unnecessary `ones()` parameter in `LorentzModel.m` and `TwoLorentzModel.m`
- Philips \*.spar file extension is no longer saved in the output structure
- Removed calculation of `df` in `GannetLoad.m`; added `dt` in its place
- Exported CSV files will no longer be overwritten if a CSV file of the same name already exists in
  the output directory; instead, a new file is created and appended with an integer
- Added new output structure attribute `nrows_water`
- Reordered structure subfields for metabolites in `GannetFit.m` and `ExportToCSV.m`
- Shifted Gannet logo and version number in PDF outputs slightly to the right
- Alpha values are now included in exported CSV files
- Replaced the Gannet logo with a PNG version
- MANY COSMETIC CHANGES

### Bug fixes

- Added a check for letter case of \*.dcm and \*.ima files in `DICOMRead.m` and `SiemensDICOMRead.m`
  (`dir()` is case-sensitive in Linux)
- Added a fix to avoid an error where MATLAB tried to close figures following an unrelated internal
  error, but no figures were created
- Fixed a bug in `NIfTIMRSRead.m` that led to an error if the `Manufacturer` field is not found in
  the data file header
- If `export_fig.m` cannot be found on the search path and `append` has been flagged in
  `GannetPreInitialise.m`, it is made clear to the user that PDFs will be saved separately
- Fixed bugs in `AlignUsingH2O.m`
- Fixed a bug in `SiemensDICOMRead.m` where an error occurred if `centerFreq` is not found in the
  data file header

## 3.3.0
<span style="color: grey;">2022-10-22</span>

### New features
- Added functionality to read NIfTI-MRS files (<a href="https://onlinelibrary.wiley.com/doi/10.1002/mrm.29418" target="_blank">Clarke et al., MRM, 2022</a>) (`NIfTIMRSRead.m`); includes additional new function `GannetMask_NIfTI.m` for voxel co-registration
- Added support for Utah Siemens MEGA-PRESS sequence (DICOM files)
- Added support for CMRR PRESS data (DICOM files)
- Added support for Siemens XA30 sequence as provided by JHU
- Changed `README` to `README.md`, which now includes badges

### Major changes
- Coil combination of GE, NIfTI-MRS, and Siemens TWIX files is now performed using generalized least squares (<a href="https://onlinelibrary.wiley.com/doi/10.1002/jmri.23941" target="_blank">An et al., JMRI, 2013</a>); this approach has been shown to result in optimal SNR and reduce spectral artifacts
- The Gannet logo has been reverted to the original image of a gannet

### Minor changes
- Added functionality for GE data in `AlignSubSpectra_PreAlignRef.m`
- An error is presented if not enough inputs are given for any of the Gannet modules
- CoRegStandAlone outputs can be hidden if `MRS_struct.p.hide` is set to `1` in `CoRegStandAlone.m` (default is `0`)
- If the user has no internet connection, `VersionCheck.m` and `UpdateGannet.m` will not run
- Updated dicm2nii to latest version (version 2022.09.15)
- Updated mapVBVD functions to latest version (<a href="https://github.com/pehses/mapVBVD" target="_blank">https://github.com/pehses/mapVBVD</a>)
- Turned off warnings of iteration limits in `SpectralRegistrationHERMES.m` and `GannetFit.m`
- Many cosmetic changes in the code

### Bug fixes
- Fixed bug in error reporting in `GannetFit.m`
- CoRegStandAlone routines now parse version numbers from the main modules properly
- Variable `filesExist` in `GannetSegment.m` may not populate; it is now pre-initialized as a precaution
- Bug fix in `GannetMask_SiemensRDA.m`
- Bug fixes for reading DICOM files

## 3.2.1
<span style="color: grey;">2022-06-07</span>

### Minor changes
- `PhaseCorrection.m` now deals better with spectra that have a non-zero baseline
- SPM12 now outputs forward deformation field (needed for normalizing voxel masks to MNI space)
- Updated `export_fig` to v3.27
- `CoRegStandAlone.m` now exports a CSV file
- Various cosmetic changes

### Bug fixes
- Fixed sub-spectra alignment for GE HERCULES data
- Fixed bug in `CoRegStandAlone.m`
- Fixed bug in `read_dcm_header.m` (thanks to Meredith Reid)

## 3.2.0
<span style="color: grey;">2021-07-30</span>

### New features

- Added `GannetVersion.m`, `VersionCheck.m`, and `UpdateGannet.m`; these new functions let users know the version of Gannet they have, if a newer release of Gannet is available, and will allow them to automatically update if wanted
- Added `ToolboxCheck.m` to check for missing MATLAB toolboxes needed to run Gannet
- Error dialog boxes are displayed at the end of the analysis pipeline if an error occurred (during a batch analysis, Gannet will skip to the next dataset and continue running)
- Users can now choose to use weighted (the default) or conventional signal averaging in `GannetPreInitialise.m` (outlier scans are removed when using conventional averaging, as before)
- Added a new (optional) method for robust, weighted signal averaging: weighted averaging by criterion function minimization (WACFM) using the generalized Cauchy distribution as the cost function; see doi:<a href="https://dx.doi.org/10.1016/j.bbe.2015.06.002" target="_blank">10.1016/j.bbe.2015.06.002</a> for more details; additionally, instead of weighting difference pairs, the subspectra of each editing subexperiment are weighted to generate weighted subspectra, which are then subtracted; see `SignalAveraging.m`
- Yair Altman's <a href="https://github.com/altmany/export_fig" target="_blank">export_fig toolbox</a> is included as an optional toolbox that, if added to MATLAB's search path, Gannet will use to append all output PDFs from each module into a single PDF (a flag must be set in `GannetPreInitialise.m` to do this); this is very useful for easy QA of lots of data; note that <a href="https://www.ghostscript.com/" target="_blank">Ghostscript</a> needs to be installed for the appending of PDFs to work (NB: macOS users should use this <a href="https://pages.uoregon.edu/koch/" target="_blank">version</a> or use <a href="https://formulae.brew.sh/formula/ghostscript" target="_blank">Homebrew</a> to install Ghostscript)
- Implemented use of hyperlinks in info and error messages printed in the command window so that users can: (1) directly run a certain function (such as `UpdateGannet.m` when `VersionCheck.m` informs the user a new release is available); (2) go straight to the line of code where an encountered error can be fixed (such as errors in `GannetPreInitialise.m`); or (3) go to a website by clicking on the printed URL (such as to download SPM12); the relevant code has been borrowed from export_fig
- Added an option in `GannetPreInitialise.m` that allows users, when using robust spectral registration, to use the averaged pre-aligned subspectra as references for aligning the averaged post-aligned subspectra (this may be helpful when robust spectral registration makes the alignment of already good-quality pre-aligned spectra worse)
- `PaperPlot.m` now allows users to plot an exemplary voxel mask co-registered to the respective structural image
- Added `GannetMask_GE_nii.m`; GE users can now use NIfTI images instead of DICOMs to co-register their MRS voxels to structural images
- Added the ability to trim datasets
- Extended the ability to concatenate files, including over multiple subjects
- Added a flag in `GannetPreInitialise.m` (`MRS_struct.p.hide`) to prevent output figures from displaying (useful when processing a large batch of files)

### Major changes

- The outdated manual PDF has been removed; up-to-date software documentation will now be found online: https://markmikkelsen.github.io/Gannet-docs/index.html
- Gannet is now licensed under the BSD 3-Clause License
- Alpha tissue correction is now performed on a by-metabolite basis rather than fixing the intrinsic WM:GM concentration to 1:2 (alpha) for all metabolites (assuming for now the ratio is 1:2 for GABA and Glx and 1:1 for GSH, EtOH and Lac); the alpha that is assumed is displayed in the GannetQuantify output figure
- Removed `MRS_struct.p.ON_OFF_order` from `GannetPreInitialise.m`; the order of editing pulses is now determined automatically; the relevant code has been borrowed from Osprey (this has only been tested to a limited extent so may require further tweaking)
- The exception to this change is phantom data, where users will still need to specify the order of editing pulses

### Minor changes

- Datasets with differing number of averages can now be batch-processed together
- Turned off the progress bar displayed when loading Siemens TWIX data to reduce loading time
- Renamed some files for the sake of clarity and style; removed obsolete `GannetMask.m`
- Signal averaging and subtraction are now done in the new function `SignalAveraging.m`
- Various improvements to `RobustSpectralRegistration.m` (formerly `Robust_Spectral_Registration.m`) for better handling of data acquired using very strong water suppression
- The assumed concentration of pure water in `GannetFit.m` and `GannetFitPhantom.m` (PureWaterConc) has been changed from 55 to 55.51 mol/kg; this puts it in line with the same constant used in `GannetQuantify.m`
- Added edits from Ralph Noeske (GE Berlin) to `GERead.m` for better handling of data when nechoes == 1; also changed the scaling factor if nechoes > 1 (hopefully this makes water-referenced measurements consistent across all data encoding flavors)
- In `GERead.m` and `PhilipsRead_data.m`, instead of using the first point in the FID for signal weighting, the most common point that is the max in the magnitude signal across all FIDs is now used
- Some changes to `SiemensDICOMRead.m` and `DICOMRead.m` for smarter handling of data files: for each dataset, metabolite and water data files should be stored in separate folders
- Improved `ExportToCSV.m`; a single .csv file is now exported for HERMES datasets; useful variables such as signal areas and Cr fit quality metrics are now also exported
- Rewrote `DiscernDataType.m` (formerly `GannetDiscernDatatype.m`)
- Minor changes and improvements to `PaperPlot.m`
- Minor changes to `CoRegStandAlone.m` functions
- Minor changes to `SignalFilter.m`
- Simplified `EddyCurrentCorrection.m` (formerly `phase_correction_fids.m`)
- Set up SPM12 for batch tissue segmentation only when tissue segmentation has not already been run and do it only once in a batch
- AllFramesFT (for the last loaded dataset) is now also saved in `MRS_struct.spec`
- When saving the `MRS_struct` structure, the .mat file is now saved using version 7.3 (for the rare case when the structure is >= 2 GB on 64-bit computers)
- Removed `MRS_struct.p.sdat` from `GannetPreInitialise.m` and the corresponding code from `GannetLoad.m` (this feature was never really used by users afaik)
- Improved global zero-order phasing in `SpectralRegistrationHERMES.m`
- Many cosmetic/aesthetic changes
  - The module output figures now show information useful for reporting acquisition parameters and results (in line with the <a href="https://onlinelibrary.wiley.com/doi/10.1002/nbm.4484" target="_blank">MRSinMRS</a> checklist)
  - Co-registered/segmented voxel masks displayed in GannetCoRegister, GannetSegment, and GannetQuantify output figures are now shown in yellow rather than grayscale
  - Smarter stacking of HERMES difference spectra in `PlotPrePostAlign.m` and `PlotPrePostAlign2.m` (formerly `GannetPlotPrePostAlign.m` and `GannetPlotPrePostAlign2.m`)
  - The Gannet documentation website URL is printed at the bottom of every output figure
  - Removed irrelevant messages from being printed in the command window during the analysis pipeline, replacing them with more useful messages
  - Minor rearranging of results text in GannetSegment and GannetQuantify output figures
  - Slight change to the y-axis labels of the Cr frequency spectrogram in the GannetLoad output figure
  - Better reordering of `MRS_struct.out` subfields created in `GannetFit.m`

### Bug fixes

- Frequency and phase offset estimates are now tracked correctly and saved when running any of the spectral registration-based alignment methods
- Removed old debugging conditional statement in `TWIXDeIdentify.m`
- Explicitly specify color of water frequency trace in GannetLoad output figure as there appears to be a bug since MATLAB R2019b when using hold to overlay plots
- Added EtOH signal parameters to `GannetQuantify.m`
- In `SiemensTwixRead.m`, if no water reference is provided, use the most common point that is the max in the magnitude signal across all FIDs instead of just using the first point
- Fixed an orientation problem in `GannetMask_GE.m`
- Removed unnecessary complex conjugate transpose during FFT of water data in `GannetLoad.m`
- Turned off warnings about legacy number generator in `RobustSpectralRegistration.m` (for when rng and randn are run in `SignalFilter.m`) if they are already turned on

## 3.1.5
<span style="color: grey;">2020-03-02</span>

### New features

- Smarter checking of incorrect settings in `GannetPreInitialise.m` by `CheckTargets.m`
- Major changes to `GannetFitPhantom.m` to bring it more in line with `GannetFit.m`
- Added support for CMRR MEGA-sLASER sequence in `SiemensTwixRead.m`
- In `SubSpectralAlign.m`, if very strong water suppression was used, use Cho instead of residual water to align subspectra of GABA-, Lac-, and EtOH-edited datasets

### Minor changes

- Noise estimation in `CalcNoise.m`, `Robust_Spectral_Registration.m`, and `SignalFilter.m` is now performed between 8 and 10 ppm to account for datasets that have shorter spectral widths
- Cleaned up code in `GannetFit.m`
- Quantification of Lac now explicitly defined as Lac+MM
- Improved functionality of `PaperPlot.m`
- In `SiemensTwixRead.m`, when a water reference is present, coil sensitivity and phase is based on the point index that is at the top of the echo, rather than simply fixing it to the first point

### Bug fixes

- Minor bug fixes in CoRegStandAlone functions
- Corrected use of random number generation in `SignalFilter.m` when performing noise replacement to ensure the same results are reproduced each time it is run on a dataset

## 3.1.4
<span style="color: grey;">2019-10-03</span>

### Bug fix

- Fixed error in the calculation of `AvgDeltaF0` for MEGA-PRESS data

## 3.1.3
<span style="color: grey;">2019-08-06</span>

### New features

- Major improvements to `Robust_Spectral_Registration.m`
- Wavelet toolbox no longer required
- Raw data is now phase-corrected prior to alignment
- `GannetLoad.m` and `GannetFit.m` no longer crash if an error occurs during a batch analysis
- An error report is printed after the final dataset is loaded/fitted so users can check which datasets failed
- Added support for GE revision header 20.007

### Minor change

- `Water_Positive` flag in `GannetPreInitialise.m` removed





